---
title: "Quant1 Homework 3"
author: "Carl Smith"
date: "11/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r  results = "asis"}
# LIBRARIES
library(tidyverse) # For readr, dplyr, ggplot2
library(countrycode) # For converting country codes between formats and adding UN region codes for proximity
library(stargazer) # For regression modeling
library(pander)

# IMPORTING AND FORMATTING DATA
# Main independent variable
# AidData Geocoded Global Chinese Official Finance, Version 1.1.1
# Utilizing transactions categorized as Official Development Flows according to OECD methodology
# Covers period 2000-2014
# https://www.aiddata.org/data/geocoded-chinese-global-official-finance-dataset
aidDataODF <- read_csv("oda-like_flows.csv") %>% 
  # Selecting relevant columns
  select(recipient_iso3, usd_current) %>% 
  # Filtering for non-zero values
  filter(usd_current != 0) %>% 
  # Filtering out regional data and keeping individual country data using three-character ISO country codes
  filter(str_length(recipient_iso3) == 3) %>% 
  # Grouping by country
  group_by(recipient_iso3) %>% 
  # Summarising each country's total aid in millions USD
  summarise(Total.CN.ODF.in.Mil = sum(usd_current/1000000))


# Dependent variable
# United Nations General Assembly Voting Data - Ideal Point table
# Voeten, Erik; Strezhnev, Anton; Bailey, Michael
# https://dataverse.harvard.edu/dataset.xhtml?persistentId=hdl:1902.1/12379
# Filtered to include only the 70th UN Session, beginning September 2015
unVoting <- read_tsv("Idealpoints.tab") %>% 
  # Selecting relevent columns
  select(session, countryname, idealpointwithpara, pctagreechina, pctagreeus) %>%
  # Naming key variables
  mutate(country.name = countryname,
    ideal.point = idealpointwithpara,
    # Converting decimals to 1-100 percentages
    pct.agree.china = pctagreechina * 100,
    pct.agree.us = pctagreeus * 100
    ) %>%
  # Filtering for 70th session beginning September 2015
  filter(session == 70)
  # Converting COW Country Codes and names to ISO 3 Character codes to facilitate merging (ugly hack to fix countrycode library not   recognising Yemen Arab Republic as Yemen)
  yemenFix <- "YEM"
  names(yemenFix) <- c("Yemen Arab Republic")
  unVoting$countryISO <- countrycode(unVoting$country.name, "country.name", "iso3c",  warn = TRUE, nomatch = NA, custom_match = yemenFix)

 
# Control variable   
# World Bank GDP per capita, PPP (current international $) by country 2014
# https://data.worldbank.org/indicator/NY.GDP.PCAP.PP.CD?most_recent_year_desc=true
gdpWorldBank <- read_csv("API_NY.GDP.PCAP.PP.CD_DS2_en_csv_v2_422115.csv") %>% 
  select('Country Code', '2014') %>%
  mutate(gdppc.2014.in.Bil = `2014`)

# Control variable
# Composite Index of Natonal Capability scores from COW
# 0-1.0 Composite of total population, urban population, iron and steel production, energy consumption, military personnel, and military expenditure
# As used by Dreher, Nunnenkamp and Thiele(2008)
# http://www.correlatesofwar.org/data-sets/national-material-capabilities
cincScore <- read_csv("NMC_5_0.csv") %>%
  filter(year == 2012) %>%
  mutate(cinc.score = `cinc`) 
cincScore$countryISO <- countrycode(cincScore$stateabb, "cowc", "iso3c",  warn = TRUE, nomatch = NA, custom_match = yemenFix)

#pander(cincScore)


# # MERGING DATA
merged <- left_join(unVoting, aidDataODF, by = c("countryISO" = "recipient_iso3"))
merged <- left_join(merged, gdpWorldBank, by = c("countryISO" = "Country Code"))
merged <- left_join(merged, cincScore, by = c("countryISO" = "countryISO"))
# Trimming unnecessary columns
merged <- merged %>%
  select(country.name, Total.CN.ODF.in.Mil, pct.agree.china, pct.agree.us, ideal.point, gdppc.2014.in.Bil, cinc.score) %>%
  # Replacing NA with 0
  mutate(Total.CN.ODF.in.Mil = replace_na(Total.CN.ODF.in.Mil, 0)) %>%
  mutate(gdppc.2014.in.Bil = replace_na(gdppc.2014.in.Bil, 0))

pander(merged)

# MODELING
stargazer(
  m1 <- lm(pct.agree.china ~ Total.CN.ODF.in.Mil, merged),
  m2 <- lm(pct.agree.china ~ Total.CN.ODF.in.Mil + ideal.point, merged),
  m3 <- lm(pct.agree.china ~ Total.CN.ODF.in.Mil + ideal.point + gdppc.2014.in.Bil, merged),
  m4 <- lm(pct.agree.china ~ Total.CN.ODF.in.Mil + ideal.point + gdppc.2014.in.Bil + cinc.score, merged),
  title = "Percent Agreement with China UN Voting 2015",
  dep.var.labels = "Percent Agreement with China UN GA Voting 2015",
  type = "html"
)